Visualizing probability distributions across bivariate cyclic temporal granularities

### 2020-07-27 main.Rmd
fixing MMPD section

\noindent The data structure suggests that we look at pairs of cyclic granularities at a time from $N_C$ cyclic granularities. Hence, all possible pairs of cyclic time granularities that could be analyzed together is $^{N_C} P_2$, with each of the circular granularity either mapped to facets or x-axis. Now, if $N_C$ is fairly large, even after excluding clashes, the list of harmonies left would be very large and overwhelming for human consumption. Hence, there is a need to rank these harmonies basis how well they capture the variation in the measured variable.<!--The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities. For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum. The harmony pair which has more variation between x-axis levels for all facet levels, on average, is likely to be a good measure of how important a harmony pair is displaying variation in the distributions. -->

@@

\noindent The choice of aesthetic mapping and facets for visualizing data sets <$C_i$, $C_j$, $v$> leads to placing $A_k$'s in close proximity with $B_l$'s representing a group/facet. A harmony pair ($C_i$, $C_j$) would be a more interesting display if the maximum pairwise distance between $A_k$'s $\forall k \in \{1, 2, \dots, K\}$ are higher for each group/facet. A probable solution for ranking interesting harmony pairs could be to arrange from highest to lowest average maximum pairwise distances. But maximum is not robust to the number of levels and seems to be higher for harmonies with higher number of levels. Thus these maximum pairwise distances need to be normalized in a way that eliminates the effect of different levels.<!--- \noindent The Fisher–Tippett–Gnedenko theorem states that the maximum of a sample of iid random variables after proper re-normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process. --> Suppose $d_{1},d_{2}\ldots ,d_{m}$ be a sequence of independent and identically-distributed pairwise distances and $M_{m}=\max\{d_{1},\ldots ,d_{m}\}$. Then Fisher–Tippett–Gnedenko theorem [@De_Haan2007-yx] suggests that if a sequence of pairs of real numbers $(a_{m}, b_{m})$ exists such that each $a_{m}>0$ and $\lim _{{m\to \infty }}P\left({\frac  {M_{m}-b_{m}}{a_{m}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, Fréchet or Weibull family. The normalizing constants $(a_{m}, b_{m})$ vary depending on the underlying distribution of the pairwise distances. Hence to normalize appropriately, it is important to assume a distribution of these distances.

@@

```{r harmony-tab, echo=FALSE, cache=TRUE}
if(!dir.exists("cache/"))
{
  dir.create("cache")
}
if (!file.exists("cache/smart_harmony.rds")) {
  sm <- smart_meter10 %>%
  filter(customer_id %in% c(10017936))

@@

smart_harmony <- sm %>%
  rank_harmony(harmony_tbl = harmonies,response =  "general_supply_kwh") %>%
  rename(`facet variable` = facet_variable,
         `x-axis variable` = x_variable,
         `facet levels` = facet_levels,
         `x-axis levels` = x_levels,
         `average maximum variation` = mean_max_variation)

  write_rds(smart_harmony,"cache/smart_harmony_ak.rds")
}
smart_harmony <- read_rds("cache/smart_harmony.rds")

>>>>

\noindent The data structure suggests that we look at pairs of cyclic granularities at a time from $N_C$ cyclic granularities. Hence, all possible pairs of cyclic time granularities that could be analyzed together is $^{N_C} P_2$, with each of the circular granularities either mapped to facets or x-axis. Now, if $N_C$ is fairly large, even after excluding clashes, the list of harmonies left would be very large and overwhelming for human consumption. Hence, there is a need to refine our search and only look at harmonies for which variation in the categories of the cyclic granularities are significant. Moreover, there is a need to rank these selected harmonies from highest to lowest variation of their categories.

<!-- rank these harmonies basis how well they capture the variation in the measured variable.<!--The harmony pairs $(C_i, C_j)$ with higher average variation in the measured variable between different levels of $C_i$ and $C_j$ has a higher chance to bring out distinct characteristics of the data.<!-- Ranking could be based on how well a harmony pair displays the variation in distributions of the measured variable within and across each levels of the cyclic granularities. For each facet level, the most interesting display is the one in which the difference in the distribution of the measured variable between the x-axis categories is maximum. The harmony pair which has more variation between x-axis levels for all facet levels, on average, is likely to be a good measure of how important a harmony pair is displaying variation in the distributions. --> -->

@@

\noindent The choice of data structure for visualizing data sets <$C_i$, $C_j$, $v$> leads to placing $A_k$'s in close proximity with $B_l$'s representing a group/facet. A harmony pair ($C_i$, $C_j$) would be a more interesting display if the maximum pairwise distance between $A_k$'s $\forall k \in \{1, 2, \dots, K\}$ are higher for each group/facet. A probable solution for ranking interesting harmony pairs could be to arrange from highest to lowest average maximum pairwise distances. But maximum is not robust to the number of levels and seems to be higher for harmonies with higher number of levels. Thus these maximum pairwise distances need to be normalized in a way that eliminates the effect of different levels.<!--- \noindent The Fisher–Tippett–Gnedenko theorem states that the maximum of a sample of iid random variables after proper re-normalization can converge in distribution to only one of Weibull, Gumbel or Freschet distribution, independent of the underlying data or process. --> Suppose $d_{1},d_{2}\ldots ,d_{m}$ be a sequence of independent and identically-distributed pairwise distances and $M_{m}=\max\{d_{1},\ldots ,d_{m}\}$. Then Fisher–Tippett–Gnedenko theorem [@De_Haan2007-yx] suggests that if a sequence of pairs of real numbers $(a_{m}, b_{m})$ exists such that each $a_{m}>0$ and $\lim _{{m\to \infty }}P\left({\frac  {M_{m}-b_{m}}{a_{m}}}\leq x\right)=F(x)$, where $F$ is a non-degenerate distribution function, then the limit distribution $F$ belongs to either the Gumbel, Fréchet or Weibull family. The normalizing constants $(a_{m}, b_{m})$ vary depending on the underlying distribution of the pairwise distances. Hence to normalize appropriately, it is important to assume a distribution of these distances.

@@

```{r harmony-tab, echo=FALSE, cache=TRUE}
# if(!dir.exists("cache/"))
# {
#   dir.create("cache")
# }
# if (!file.exists("cache/smart_harmony.rds")) {
  sm <- smart_meter10 %>%
  filter(customer_id %in% c(10017936))

@@

smart_harmony <- sm %>%
  rank_harmony(harmony_tbl = harmonies,response =  "general_supply_kwh") %>%
  rename(`facet variable` = facet_variable,
         `x-axis variable` = x_variable,
         `facet levels` = facet_levels,
         `x-axis levels` = x_levels,
         `average maximum variation` = MMPD)

#   write_rds(smart_harmony,"cache/smart_harmony_ak.rds")
# }
#smart_harmony <- read_rds("cache/smart_harmony.rds")
